### 2022-02-28 doc/src/MSLevel.tex

%=======================================================================
\chapter{Interrupts for Machine and Supervisor Levels}
\label{ch:MSLevel}

\textbf{%
Warning!
This draft specification is likely to change before being accepted as
standard by the {\RISCV} International Association.%
}
\bigskip

The {\RISCV} Privileged Architecture defines several major identities
in the range 0--15 for interrupts at a hart, including machine-level
and supervisor-level external interrupts (numbers 11 and~9), machine-
and supervisor-level timer interrupts (7~and~5), and machine- and
supervisor-level software interrupts (3~and~1).
Beyond these major labels, the \emph{external} interrupts at each
privilege level are given secondary, minor identities by an external
interrupt controller such as an APLIC or IMSIC, distinguishing interrupts
from different devices or causes.
These minor identities for external interrupts were covered in Chapters
\ref{ch:IMSIC} and~\ref{ch:AdvPLIC} specifying the IMSIC and
APLIC components.

The Advanced Interrupt Architecture assigns other major interrupt
identities in the range 16--63 to additional \emph{local interrupts}
that arise within or in close proximity to the hart, often for
reporting errors.
A mechanism is also defined that allows software to
selectively delegate both local and custom interrupts to the next lower
privilege level, or in some cases to inject entirely virtual interrupts
into a lower privilege level.

Lastly, an optional facility lets software assign priorities to major
interrupts (such as the timer and software interrupts, and any local
interrupts) such that they may mix with the priorities set for external
interrupts by a PLIC, APLIC, or IMSIC.

%-----------------------------------------------------------------------
\section{Interrupt allocations and default priorities}
\label{sec:majorIntrs}

The base {\RISCV} Privileged Architecture designates major interrupt
numbers 16 and above as being for platform or custom use.
The Advanced Interrupt Architecture allocates 24 of these interrupt
numbers to standardize as local interrupts, and leaves the rest for
custom use.

\begin{table*}[h!]
\begin{center}
\begin{tabular}{|c|l|l|}
\hline
Default        &                         & \\
priority order & Major interrupt numbers & \multicolumn{1}{c|}{Description} \\
\hline
\hline
Highest & 63, 62, 31, 30, 61, 60, & Even numbers are standard local
                                     interrupts \\
        & 59, 58, 29, 28, 57, 56, & \ or reserved; odd numbers are
                                     designated \\
        & 55, 54, 27, 26, 53, 52, & \ for custom use \\
        & 51, 50, 25, 24, 49, 48  & \\
\cline{2-3}
        & 11, 3, 7   & Machine interrupts:  external, software, timer \\
        & \ 9, 1, 5  & Supervisor interrupts:  external, software, timer \\
        & 12         & Supervisor guest external interrupt \\
        & 10, 2, 6   & VS interrupts:  external, software, timer \\
\cline{2-3}
        & 47, 46, 23, 22, 45, 44, & Even numbers are standard local
                                     interrupts \\
        & 43, 42, 21, 20, 41, 40, & \ or reserved; odd numbers are
                                     designated \\
        & 39, 38, 19, 18, 37, 36, & \ for custom use \\
Lowest  & 35, 34, 17, 16, 33, 32  & \\
\hline
\end{tabular}
\end{center}
\caption{%
The default priority order for major interrupts 0--63.
Interrupt~63 has the highest default priority, and interrupt~32 has the
lowest default priority.%
}
\label{tab:defaultIntrPrios}
\end{table*}

Table~\ref{tab:defaultIntrPrios} shows the allocation of major
interrupt numbers 0--63, listed in default priority order from highest
to lowest.
Note that, for interrupts with major identities in the range 16--31, a
higher interrupt number conveys higher default priority, and likewise
for major identities in the range 32--63.
These two groups are interleaved together in the complete order, and
the Privileged Architecture's standard interrupts, 0--15, are inserted
into the middle of the sequence.

\begin{commentary}
The default priority order is arranged so that interrupts 0--31 can
potentially be an adequate subset on their own for \mbox{32-bit}
{\RISCV} systems.
\end{commentary}

\begin{commentary}
If the {\RISCV} Privileged Architecture defines interrupt~0, the
Advanced Interrupt Architecture needs it to have a default priority
lower than certain external interrupts.
See Sections \ref{sec:mtopi} and~\ref{sec:stopi} on CSRs \z{mtopi} and
\z{stopi}.
\end{commentary}

Interrupt numbers 64 and above are all designated for custom use.
If a hart implements custom interrupts with these numbers, their
positions in the default priority order must be documented for the
hart.
Default priority order is standardized by the Advanced Interrupt
Architecture only for interrupts 0--63.

When a hart supports the arbitrary configuration of interrupt
priorities by software (described in later sections), the default
priority order still remains relevant for breaking ties when two
interrupt sources are assigned the same priority number.

%-----------------------------------------------------------------------
\section{Defined local interrupts}

The local interrupts currently defined by the Advanced Interrupt
Architecture are listed here, in order of decreasing default
priority:\nopagebreak
\begin{displayLinesTable}[l@{\quad}l]
30 & Reserved for standard reporting of bus or system errors \\
58 & Tentatively reserved for per-core high-power/over-temperature events \\
54 & Reserved for standard high-priority RAS events \\
\noalign{\bigskip}
38 & Reserved for standard low-priority RAS events \\
18 & Debug/trace interrupt \\
\end{displayLinesTable}
\emph{RAS} is an abbreviation for \emph{Reliability, Availability, and
Serviceability}.

\begin{commentary}
The Advanced Interrupt Architecture reserves local interrupt numbers
for use by other {\RISCV} standards.
Complete facilities for managing bus or system errors,
high-power/over-temperature events, RAS events, or debug/trace
interrupts are not defined here.
\end{commentary}

The Advanced Interrupt Architecture does not itself require that all
local interrupts it defines be implemented.
A {\RISCV} hart may implement any subset of them, or none of them.
For whichever ones are implemented, the corresponding bits in CSRs
\z{mip}/\z{miph} and \z{mie}/\z{mieh} must be writable, and the
corresponding bits in \z{mideleg}/\z{midelegh} (if those CSRs exist
because supervisor mode is implemented) may each either be writable or
be hardwired to zero.

Even when a defined local interrupt is implemented, it is not
necessarily the case that all events matching the description for that
interrupt actually cause the local interrupt to be raised.
For example, a single {\RISCV} system might report bus errors in
multiple ways, depending on where in the system the error is detected.
Some bus errors might signal the standard local interrupt (number~30),
while others are reported as external interrupts, routed through an APLIC
or IMSIC.

%-----------------------------------------------------------------------
\section{Interrupts at machine level}

When a hart implements one of the local interrupts defined by the
Advanced Interrupt Architecture (an even interrupt number in the
range 16--62), the corresponding bit position is writable in both
\z{mip}/\z{miph} and \z{mie}/\z{mieh}.
An occurrence of the interrupt event causes the interrupt-pending bit
in \z{mip}/\z{miph} to be set to one.
This bit then remains set until cleared by software.

As established by the {\RISCV} Privileged Architecture, an interrupt
traps to \mbox{M-mode} whenever all of the following are true:
(a)~either the current privilege mode is \mbox{M-mode} and
machine-level interrupts are enabled by the MIE bit of \z{mstatus}, or
the current privilege mode has less privilege than \mbox{M-mode};
(b)~matching bits in \z{mip}/\z{miph} and \z{mie}/\z{mieh} are both
one; and
(c)~if \z{mideleg} exists, the corresponding bit in
\z{mideleg}/\z{midelegh} is zero.

When multiple interrupt causes are ready to trigger simultaneously, the
interrupt taken first is determined by priority order, which may be the
default order specified earlier, or may be a modified order configured
by software.

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
\subsection{Configuring priorities of major interrupts at machine level}
\label{sec:intrPrios-M}

The machine-level priorities for major interrupts 0--63 may be
configured by a set of registers accessed through the \z{miselect} and
\z{mireg} CSRs introduced in Chapter~\ref{ch:CSRs}.
When MXLEN =~32, sixteen of these registers are defined, listed below
with their \z{miselect} addresses:
\begin{displayLinesTable}[c@{\quad}l]
\z{0x30} & \z{iprio0} \\
\z{0x31} & \z{iprio1} \\
\dots    & \ \dots \\
\z{0x3F} & \z{iprio15} \\
\end{displayLinesTable}
Each register controls the priorities of four interrupts, with one
\mbox{8-bit} byte per interrupt.
For a number~$k$ in the range 0--15, register \z{iprio}$k$
controls the priorities of interrupts $k\times\mbox{4}$ through
${k\times\mbox{4}+\mbox{3}}$, formatted as follows:
\begin{displayLinesTable}[l@{\quad}l]
bits 7:0   & Priority number for interrupt $k\times\mbox{4}$ \\
bits 15:8  & Priority number for interrupt $k\times\mbox{4}+\mbox{1}$ \\
bits 23:16 & Priority number for interrupt $k\times\mbox{4}+\mbox{2}$ \\
bits 31:24 & Priority number for interrupt $k\times\mbox{4}+\mbox{3}$ \\
\end{displayLinesTable}

When MXLEN =~64, only the even-numbered registers exist:\nopagebreak
\begin{displayLinesTable}[c@{\quad}l]
\z{0x30} & \z{iprio0} \\
\z{0x32} & \z{iprio2} \\
\dots    & \ \dots \\
\z{0x3E} & \z{iprio14} \\
\end{displayLinesTable}
Each register controls the priorities of eight interrupts.
For even~$k$ in the range 0--14, register \z{iprio}$k$
controls the priorities of interrupts $k\times\mbox{4}$ through
${k\times\mbox{4}+\mbox{7}}$, formatted as follows:
\begin{displayLinesTable}[l@{\quad}l]
bits 7:0   & Priority number for interrupt $k\times\mbox{4}$ \\
bits 15:8  & Priority number for interrupt $k\times\mbox{4}+\mbox{1}$ \\
bits 23:16 & Priority number for interrupt $k\times\mbox{4}+\mbox{2}$ \\
bits 31:24 & Priority number for interrupt $k\times\mbox{4}+\mbox{3}$ \\
bits 39:32 & Priority number for interrupt $k\times\mbox{4}+\mbox{4}$ \\
bits 47:40 & Priority number for interrupt $k\times\mbox{4}+\mbox{5}$ \\
bits 55:48 & Priority number for interrupt $k\times\mbox{4}+\mbox{6}$ \\
bits 63:56 & Priority number for interrupt $k\times\mbox{4}+\mbox{7}$ \\
\end{displayLinesTable}

When MXLEN =~64 and \z{miselect} is an odd value in the range
\z{0x31}--\z{0x3F}, attempting to access \z{mireg} raises an illegal
instruction exception.

The valid registers \z{iprio0}--\z{iprio15} are known collectively as
the \emph{\texttt{iprio} array} for machine level.

The width of priority numbers for external interrupts is
\emph{IPRIOLEN}.
This parameter is affected by the main external interrupt controller
for the hart, whether a PLIC, APLIC, or IMSIC.

For an APLIC, IPRIOLEN is in the range 1--8 as specified in
Chapter~\ref{ch:AdvPLIC} on the APLIC.

For an IMSIC, IPRIOLEN is 6, 7, or~8.
IPRIOLEN may be 6 only if the number of external interrupt identities
implemented by the IMSIC is~63.
IPRIOLEN may be 7 only if the number of external interrupt identities
implemented by the IMSIC is no more than 127.
IPRIOLEN may be 8 for any IMSIC, regardless of the number of external
interrupt identities implemented.

Each byte of a valid \z{iprio}$k$ register is either a read-only zero
or a {\WARL} unsigned integer field implementing exactly IPRIOLEN bits.
For a given interrupt number, if the corresponding bit in
\z{mie}/\z{mieh} is read-only zero, then the interrupt's priority
number in the \z{iprio} array must be read-only zero as well.
The priority number for a machine-level external interrupt (bits 31:24
of register \z{iprio2}) must also be read-only zero.
Aside from these two restrictions, implementations may freely choose
which priority number fields are settable and which are read-only
zeros.
If all bytes in the \z{iprio} array are read-only zeros, priorities
can be configured only for external interrupts, not for any other
interrupts.

\begin{commentary}
Platform standards may require that priorities be configurable for
certain interrupt causes.
\end{commentary}

The \z{iprio} array accessed via \z{miselect} and \z{mireg} affects the
prioritization of interrupts only when they trap to \mbox{M-mode}.
When an interrupt's priority number in the array is zero (either
read-only zero or set to zero), its priority is the default order from
Section~\ref{sec:majorIntrs}.
Setting an interrupt's priority number instead to a nonzero value~$p$
gives that interrupt nominally the same priority as a machine-level
external interrupt with priority number~$p$.
For a major interrupt that defaults to a higher priority than machine
external interrupts, setting its priority number to a nonzero value
\emph{lowers} its priority.
For a major interrupt that defaults to a lower priority than machine
external interrupts, setting its priority number to a nonzero value
\emph{raises} its priority.
When two interrupt causes have been assigned the same nominal priority,
ties are broken by the default priority order.
Table~\ref{tab:intrPrios-M} summarizes the effect of priority numbers
on interrupt priority.

\begin{table*}[h!]
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
 & Interrupts with default & Machine external & Interrupts with default \\
 & priority above machine  & interrupts       & priority below machine \\
 & external interrupts     &                  & external interrupts \\
\hline
         & Priority number in & Priority number from & Priority number in \\
Priority & machine-level      & interrupt controller & machine-level \\
order    & \z{iprio} array    & (APLIC or IMSIC)     & \z{iprio} array \\
\hline
\hline
Highest  & 0                  &                      & \\
\cline{2-4}
         & 1                  & 1                    & 1 \\
         & 2                  & 2                    & 2 \\
         & $\cdots$           & $\cdots$             & $\cdots$ \\
         & 254                & 254                  & 254 \\
         & 255                & 255                  & 255 \\
\cline{2-4}
         &                    & 256 and above        & \\
         &                    & (IMSIC only)         & \\
\cline{2-4}
Lowest   &                    &                      & 0 \\
\hline
\end{tabular}
\end{center}
\caption{%
Effect of the machine-level \z{iprio} array on the priorities of
interrupts taken in \mbox{M-mode}.
For interrupts with the same priority number, the default order of
Section~\ref{sec:majorIntrs} prevails.%
}
\label{tab:intrPrios-M}
\end{table*}

\begin{commentary}
When a hart has an IMSIC supporting more than 255 minor identities
for external interrupts, the only non-default priorities that can be
configured for other interrupts are those corresponding to external
interrupt identities 1--255, not those of identities 256 or higher.
\end{commentary}

If supported, setting the priority number for supervisor-level external
interrupts (bits 15:8 of \z{iprio2}) to a nonzero value~$p$ has the
effect of giving the entire category of supervisor external interrupts
nominally the same priority as a machine external interrupt with
priority number~$p$.
But note that this applies only to the case when supervisor external
interrupts trap to \mbox{M-mode}.

(Because supervisor guest external interrupts and VS-level external
interrupts are required to be delegated to supervisor level when the
hypervisor extension is implemented, the machine-level priority numbers
for these interrupts are always ignored and should be read-only zeros.)

If the system has an original PLIC or \mbox{Duo-PLIC} for backward
compatibility with older software, reset should initialize the
machine-level \z{iprio} array to all zeros.

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
\subsection{Machine top interrupt CSR (\zSafe{mtopi})}
\label{sec:mtopi}

Machine-level CSR \z{mtopi} is read-only with width MXLEN.
A read of \z{mtopi} returns information about the highest-priority
pending-and-enabled interrupt for machine level, in this
format:\nopagebreak
\begin{displayLinesTable}[l@{\quad}l]
bits 27:16 & IID \\
bits 7:0   & IPRIO \\
\end{displayLinesTable}
All other bits of \z{mtopi} are reserved and read as zeros.

The value of \z{mtopi} is zero unless there is an interrupt pending in
\z{mip}/\z{miph} and enabled in \z{mie}/\z{mieh} that is not delegated
to a lower privilege level.
When there is a pending-and-enabled major interrupt for machine
level, field IID (Interrupt Identity) is the major identity number of
the highest-priority interrupt, and field IPRIO indicates its priority.

If all bytes of the machine-level \z{iprio} array are read-only zeros,
a simplified implementation of field IPRIO is allowed in which its
value is always 1 whenever \z{mtopi} is not zero.

Otherwise, when \z{mtopi} is not zero, if the priority number for the reported
interrupt is in the range 1 to 255, IPRIO is simply that number.
If the interrupt's priority number is zero or greater than 255,
IPRIO is set to either 0 or 255 as follows:
\begin{itemize}

\item
If the interrupt's priority number is greater than 255, then
IPRIO is 255 (lowest priority).

\item
If the interrupt's priority number is zero and interrupt number IID has
a default priority higher than a machine external interrupt, then IPRIO
is~0 (highest priority).

\item
If the interrupt's priority number is zero and interrupt number IID has
a default priority lower than a machine external interrupt, then IPRIO
is 255 (lowest priority).

\end{itemize}

\begin{commentary}
To ensure that \z{mtopi} is never zero when an interrupt is pending
and enabled for machine level, if major interrupt~0 can trap to
\hbox{M-mode}, it must have a default priority lower than a machine
external interrupt.
\end{commentary}

The {\RISCV} Privileged Architecture ensures that, when the value
of \z{mtopi} is not zero, a trap is taken to \mbox{M-mode} for the
interrupt indicated by field IID if either the current privilege
mode is~M and the MIE bit in CSR \z{mstatus} is set, or the current
privilege mode has less privilege than \mbox{M-mode}.
The trap itself does not cause the value of \z{mtopi} to change.

The following pseudocode shows how a machine-level trap handler might
read \z{mtopi} to avoid redundant restoring and saving of processor
registers when an interrupt arrives during the handling of another trap
(either a synchronous exception or an earlier interrupt):
\begin{displayLinesTable}
save processor registers \\
\z{i = }read CSR \z{mcause} \\
\z{if (i >= 0) \LB} \\
\qquad handle synchronous exception \z{i} \\
\qquad restore \z{mstatus} if necessary \\
\z{\RB}\\
\z{if (mstatus.MPIE == 1) \LB} \\
\qquad loop \z{\LB} \\
\qquad\qquad \z{i = }read CSR \z{mtopi} \\
\qquad\qquad \z{if (i == 0)} exit loop \\
\qquad\qquad \z{i = i>>16} \\
\qquad\qquad call the interrupt handler for major interrupt \z{i} \\
\qquad \z{\RB} \\
\z{\RB} \\
restore processor registers \\
return from trap \\
\end{displayLinesTable}
(This example can be further optimized, but with an increase in
complexity.)

%-----------------------------------------------------------------------
\section{Interrupt filtering and virtual interrupts for supervisor level}
\label{sec:virtIntrs-S}

When supervisor mode is implemented, the Advanced Interrupt
Architecture adds a facility for software filtering
of interrupts and for virtual interrupts, making use of new
CSRs \z{mvien}/\z{mvienh} (Machine Virtual Interrupt Enables) and
\z{mvip}/\z{mviph} (Machine Virtual Interrupt-Pending bits).
\emph{Interrupt filtering} permits a local or custom interrupt
to trap to \mbox{M-mode} and then be selectively delegated by
software to supervisor level, even while the corresponding bit in
\z{mideleg}/\z{midelegh} remains zero.
The same hardware may also, under the right circumstances, allow
machine level to assert \emph{virtual interrupts} to supervisor level
that have no connection to any real interrupt events.

Just as with CSRs \z{mip}/\z{miph}, \z{mie}/\z{mieh}, and
\z{mideleg}/\z{midelegh}, each bit of registers \z{mvien}/\z{mvienh}
and \z{mvip}/\z{mviph} corresponds with an interrupt number in the
range 0--63.
When a bit in \z{mideleg}/\z{midelegh} is zero and the matching
bit in \z{mvien}/\z{mvienh} is one, then the same bit position
in \z{sip}/\z{siph} is an alias for the corresponding bit in
\z{mvip}/\z{mviph}.
A bit in \z{sip}/\z{siph} is read-only zero when the corresponding bits
in \z{mideleg}/\z{midelegh} and \z{mvien}/\z{mvienh} are both zero.
The combined effects of \z{mideleg}/\z{midelegh} and
\z{mvien}/\z{mvienh} on \z{sip}/\z{siph} and \z{sie}/\z{sieh} are
summarized in Table~\ref{tab:intrFilteringForS}.

\begin{table*}[h!]
\begin{center}
\begin{tabular}{|c|c||c|c|}
\hline
\z{mideleg[}$n$\z{]} & \z{mvien[}$n$\z{]} &
    \z{sip[}$n$\z{]} & \z{sie[}$n$\z{]} \\
\hline
\hline
0 & 0  & Read-only 0                & Read-only 0 \\
0 & 1  & Alias of \z{mvip[}$n$\z{]} & Writable \\
1 & -- & Alias of \z{mip[}$n$\z{]}  & Alias of \z{mie[}$n$\z{]} \\
\hline
\end{tabular}
\end{center}
\caption{%
The effects of \z{mideleg} and \z{mvien} on \z{sip} and \z{sie} (except
for the hypervisor extension's VS-level interrupts, which appear in
\z{hip} and \z{hie} instead of \z{sip} and \z{sie}).
A bit in \z{mvien} can be set to~1 only for major interrupts 13--63.
When interrupts 0--12 are delegated by \z{mideleg}, the
aliases in \z{sip} may be read-only, as specified by the
{\RISCV} Privileged Architecture.%
}
\label{tab:intrFilteringForS}
\end{table*}

\begin{commentary}
The name of CSR \z{mvien} is not ``\/\z{mvie}'' because the function of
this register is more analogous to \z{mcounteren} than to \z{mie}.
The bits of\/ \z{mvien} control whether the virtual interrupt-pending
bits in register \z{mvip} are active and visible at supervisor level.
This is different than how the usual interrupt-enable bits (such as
in \z{mie}) mask pending interrupts.
\end{commentary}

A bit in \z{sie}/\z{sieh} is writable if and only if the corresponding
bit is set in either \z{mideleg}/\z{midelegh} or \z{mvien}/\z{mvienh}.
When an interrupt is delegated by \z{mideleg}/\z{midelegh}, the
writable bit in \z{sie}/\z{sieh} is an alias of the corresponding bit
in \z{mie}/\z{mieh};
else it is an independent writable bit.
As usual, bits that are not writable in \z{sie}/\z{sieh} must be
read-only zeros.

For interrupt numbers 13--63, implementations may freely choose which
bits of \z{mvien}/\z{mvienh} are writable and which bits are read-only
zero or one.
If such a bit in \z{mvien}/\z{mvienh} is read-only zero (preventing the
virtual interrupt from being enabled), the same bit should be read-only
zero in \z{mvip}/\z{mviph}.
All other bits for interrupts 13--63 must be writable in
\z{mvip}/\z{mviph}.

\begin{commentary}
Platform standards may require that interrupt filtering be supported
for certain interrupt causes.
\end{commentary}

Bits 12:0 in \z{mvien} are reserved and must be read-only zeros.
Bits 1 and~5 of \z{mvip} are aliases of the same bits in \z{mip}
(SSIP and STIP) when those bits are writable in \z{mip}.
When STIP is not writable in \z{mip} (such as when optional
CSR \z{stimecmp} is implemented), the
same bit in \z{mvip} is read-only zero.
Bit~9 of \z{mvip} is an alias of the software-writable bit~9 of \z{mip}
(SEIP).
The other bits of \z{mvip} in the range 12:0 are reserved and must be
read-only zeros.

\begin{commentary}
Bit~9 of\/ \z{mvip} makes the software-writable SEIP bit of\/ \z{mip}
directly accessible by itself.
\end{commentary}

When supervisor mode is implemented, the minimal required
implementation of \z{mvien}/\z{mvienh} and \z{mvip}/\z{mviph} has all
bits being read-only zeros except for \z{mvip} bits 1, 5, and~9, which
are aliases of existing writable bits in \z{mip}.
When supervisor mode is not implemented, registers \z{mvien}/\z{mvienh}
and \z{mvip}/\z{mviph} do not exist.

%-----------------------------------------------------------------------
\section{Interrupts at supervisor level}
\label{sec:intrs-S}

If a local interrupt defined by the Advanced Interrupt Architecture
becomes pending (=~1) in \z{sip}/\z{siph}, the bit in \z{sip}/\z{siph}
is writable and will remain set until cleared by software.

Just as for machine level, the taking of interrupt traps at supervisor
level remains essentially the same as specified by the {\RISCV}
Privileged Architecture.
An interrupt traps into \mbox{S-mode} (or \mbox{HS-mode}) whenever all
of the following are true:
(a)~either the current privilege mode is \mbox{S-mode} and
supervisor-level interrupts are enabled by the SIE bit of \z{sstatus},
or the current privilege mode has less privilege than \mbox{S-mode};
(b)~matching bits in \z{sip}/\z{siph} and \z{sie}/\z{sieh} are both
one, or, if the hypervisor extension is implemented, matching bits in
\z{hip} and \z{hie} are both one; and
(c)~if the hypervisor extension is implemented, the corresponding bit
in \z{hideleg}/\z{hidelegh} is zero.

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
\subsection{Configuring priorities of major interrupts at supervisor level}
\label{sec:intrPrios-S}

Supervisor-level priorities for major interrupts 0--63 are configurable
in an array of supervisor-level \z{iprio}$k$ registers accessed through
\z{siselect} and \z{sireg}.
This array has the same structure when SXLEN = 32 or 64 as does the
machine-level \z{iprio} array when MXLEN = 32 or 64, respectively.
To summarize, when SXLEN =~32, there are sixteen \mbox{32-bit}
registers with these \z{siselect} addresses:
\begin{displayLinesTable}[c@{\quad}l]
\z{0x30} & \z{iprio0} \\
\z{0x31} & \z{iprio1} \\
\dots    & \ \dots \\
\z{0x3F} & \z{iprio15} \\
\end{displayLinesTable}
Each register controls the priorities of four interrupts, one
\mbox{8-bit} byte per interrupt.
When SXLEN =~64, only the even-numbered registers exist:
\begin{displayLinesTable}[c@{\quad}l]
\z{0x30} & \z{iprio0} \\
\z{0x32} & \z{iprio2} \\
\dots    & \ \dots \\
\z{0x3E} & \z{iprio14} \\
\end{displayLinesTable}
Each register controls the priorities of eight interrupts.
If SXLEN = 64 and \z{siselect} is an odd value in the range
\z{0x31}--\z{0x3F}, attempting to access \z{sireg} raises an illegal
instruction exception.

The valid registers \z{iprio0}--\z{iprio15} are known collectively as
the \emph{\texttt{iprio} array} for supervisor level.
Each byte of a valid \z{iprio}$k$ register is either a read-only zero
or a {\WARL} unsigned integer field implementing exactly IPRIOLEN bits.

For a given interrupt number, if the corresponding bit in
\z{sie}/\z{sieh} is read-only zero, then the interrupt's priority
number in the supervisor-level \z{iprio} array must be read-only zero
as well.
The priority number for a supervisor-level external interrupt
(bits 15:8 of \z{iprio2}) must also be read-only zero.
Aside from these two restrictions, implementations may freely choose
which priority number fields are settable and which are read-only
zeros.

\begin{commentary}
As always, platform standards may require that priorities be
configurable for certain interrupt causes.
\end{commentary}

The supervisor-level \z{iprio} array accessed via \z{siselect} and
\z{sireg} affects the prioritization of interrupts only when they trap
to \mbox{S-mode}.
When an interrupt's priority number in the array is zero (either
read-only zero or set to zero), its priority is the default order from
Section~\ref{sec:majorIntrs}.
Setting an interrupt's priority number instead to a nonzero value~$p$
gives that interrupt nominally the same priority as a supervisor-level
external interrupt with priority number~$p$.
For an interrupt that defaults to a higher priority than supervisor
external interrupts, setting its priority number to a nonzero value
lowers its priority.
For an interrupt that defaults to a lower priority than supervisor
external interrupts, setting its priority number to a nonzero value
raises its priority.
When two interrupt causes have been assigned the same nominal priority,
ties are broken by the default priority order.
Table~\ref{tab:intrPrios-S} summarizes the effect of priority numbers
on interrupt priority.

\begin{table*}[h!]
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
 & Interrupts with default  & Supervisor external & Interrupts with default \\
 & priority above supervisor & interrupts       & priority below supervisor \\
 & external interrupts       &                  & external interrupts \\
\hline
         & Priority number in & Priority number from & Priority number in \\
Priority & supervisor-level   & interrupt controller & supervisor-level \\
order    & \z{iprio} array    & (APLIC or IMSIC)     & \z{iprio} array \\
\hline
\hline
Highest  & 0                  &                      & \\
\cline{2-4}
         & 1                  & 1                    & 1 \\
         & 2                  & 2                    & 2 \\
         & $\cdots$           & $\cdots$             & $\cdots$ \\
         & 254                & 254                  & 254 \\
         & 255                & 255                  & 255 \\
\cline{2-4}
         &                    & 256 and above        & \\
         &                    & (IMSIC only)         & \\
\cline{2-4}
Lowest   &                    &                      & 0 \\
\hline
\end{tabular}
\end{center}
\caption{%
Effect of the supervisor-level \z{iprio} array on the priorities of
interrupts taken in \mbox{S-mode}.
For interrupts with the same priority number, the default order of
Section~\ref{sec:majorIntrs} prevails.%
}
\label{tab:intrPrios-S}
\end{table*}

If supported, setting the priority number for VS-level external
interrupts (bits 23:16 of \z{iprio2}) to a nonzero value~$p$ has
the effect of giving the entire category of VS external interrupts
nominally the same priority as a supervisor external interrupt with
priority number~$p$, when VS external interrupts trap to \mbox{S-mode}.

If the system has an original PLIC or \mbox{Duo-PLIC} for backward
compatibility with older software, reset should initialize the
supervisor-level \z{iprio} array to all zeros.

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
\subsection{Supervisor top interrupt CSR (\zSafe{stopi})}
\label{sec:stopi}

Supervisor-level CSR \z{stopi} is read-only with width SXLEN.
A read of \z{stopi} returns information about the highest-priority
pending-and-enabled interrupt for supervisor level, in this
format:\nopagebreak
\begin{displayLinesTable}[l@{\quad}l]
bits 27:16 & IID \\
bits 7:0   & IPRIO \\
\end{displayLinesTable}
All other bits of \z{stopi} are reserved and read as zeros.

The value of \z{stopi} is zero unless:
(a)~there is an interrupt that is both pending in \z{sip}/\z{siph}
and enabled in \z{sie}/\z{sieh}, or, if the hypervisor extension is
implemented, both pending in \z{hip} and enabled in \z{hie}; and
(b)~if CSR \z{sideleg} or \z{hideleg} exists, the interrupt is not
delegated to a lower privilege level.
When there is a pending-and-enabled major interrupt for supervisor
level, field IID is the major identity number of the highest-priority
interrupt, and field IPRIO indicates its priority.

If all bytes of the supervisor-level \z{iprio} array are read-only
zeros, a simplified implementation of field IPRIO is allowed in which
its value is always 1 whenever \z{stopi} is not zero.

Otherwise, when \z{stopi} is not zero, if the priority number for the reported
interrupt is in the range 1 to 255, IPRIO is simply that number.
If the interrupt's priority number is zero or greater than 255,
IPRIO is set to either 0 or 255 as follows:
\begin{itemize}

\item
If the interrupt's priority number is greater than 255, then
IPRIO is 255 (lowest priority).

\item
If the interrupt's priority number is zero and interrupt number IID has
a default priority higher than a supervisor external interrupt, then
IPRIO is~0 (highest priority).

\item
If the interrupt's priority number is zero and interrupt number IID
has a default priority lower than a supervisor external interrupt, then
IPRIO is 255 (lowest priority).

\end{itemize}

\begin{commentary}
To ensure that \z{stopi} is never zero when an interrupt is pending
and enabled for supervisor level, if major interrupt~0 can trap to
\hbox{S-mode}, it must have a default priority lower than a supervisor
external interrupt.
\end{commentary}

The {\RISCV} Privileged Architecture ensures that, when the value
of \z{stopi} is not zero, a trap is taken to \mbox{S-mode} for the
interrupt indicated by field IID if either the current privilege
mode is~S and the SIE bit in CSR \z{sstatus} is set, or the current
privilege mode has less privilege than \mbox{S-mode}.
The trap itself does not cause the value of \z{stopi} to change.

The following pseudocode shows how a supervisor-level trap handler
might read \z{stopi} to avoid redundant restoring and saving of
processor registers when an interrupt arrives during the handling of
another trap (either a synchronous exception or an earlier interrupt):
\begin{displayLinesTable}
save processor registers \\
\z{i = }read CSR \z{scause} \\
\z{if (i >= 0) \LB} \\
\qquad handle synchronous exception \z{i} \\
\qquad restore \z{sstatus} if necessary \\
\z{\RB}\\
\z{if (sstatus.SPIE == 1) \LB} \\
\qquad loop \z{\LB} \\
\qquad\qquad \z{i = }read CSR \z{stopi} \\
\qquad\qquad \z{if (i == 0)} exit loop \\
\qquad\qquad \z{i = i>>16} \\
\qquad\qquad call the interrupt handler for major interrupt \z{i} \\
\qquad \z{\RB} \\
\z{\RB} \\
restore processor registers \\
return from trap \\
\end{displayLinesTable}
(This example can be further optimized, but with an increase in
complexity.)