### 2022-08-15 doc/src/preface.tex

%=======================================================================
\chapter{Preface}

This document describes an Advanced Interrupt Architecture
being proposed for {\RISCV} systems.

No part of this document has yet been ratified
by the {\RISCV} International Association.
The table below shows the current status of each chapter,
and also indicates which chapters specify extensions to the
{\RISCV} ISA (instruction set architecture) and which are non-ISA.

{
\begin{table}[hbt]
\centering
\begin{tabular}{|l|c|c|}
\hline
Chapter                                                  & ISA? & Status \\
\hline
\hline
1.\ Introduction                                         & ---  & Stable \\
2.\ Control and Status Registers (CSRs) Added to Harts   & Yes  & Stable \\
3.\ Incoming MSI Controller (IMSIC)                      & Yes  & Stable \\
4.\ Advanced Platform-Level Interrupt Controller (APLIC) & No   & Stable \\
5.\ Duo-PLIC                                             & No   & Draft \\
6.\ Interrupts for Machine and Supervisor Levels         & Yes  & Stable \\
7.\ Interrupts for Virtual Machines (VS Level)           & Yes  & Stable \\
8.\ Interprocessor Interrupts (IPIs)                     & No   & Stable \\
9.\ IOMMU Support for MSIs to Virtual Machines           & No   & Draft \\
\hline
\end{tabular}
\end{table}
}

The chapters marked in the table as \emph{Stable} are not expected
to change significantly before being put up for ratification,
but changes of all kinds are still possible nonetheless.
The chapters marked \emph{Draft} are very much
at risk of changing before ratification.
An implementation adhering to the current document might not conform
to an eventual ratified Advanced Interrupt Architecture for {\RISCV}.

Concerning the two chapters that remain in draft state:
\begin{itemize}

\item
The authors are not aware that any implementation of
a Duo-PLIC yet exists, even one just for simulation.
So long as that remains true, the Duo-PLIC specification of
Chapter~\ref{ch:DuoPLIC} will not advance beyond draft status.

\item
A separate standard is being developed for IOMMUs in {\RISCV} systems.
Chapter~\ref{ch:IOMMU} about an IOMMU's support for MSIs
(message-signaled interrupts) is expected to be finalized
only in conjunction with the full {\RISCV} IOMMU specification.  
\end{itemize}