### 2023-02-28 doc/src/preface.tex

%=======================================================================
\chapter{Preface}

This document describes an Advanced Interrupt Architecture
being proposed for {\RISCV} systems.

No part of this document has yet been ratified
by the {\RISCV} International Association.
The table below shows the current status of each chapter,
and also indicates which chapters specify extensions to the
{\RISCV} ISA (instruction set architecture) and which are non-ISA.

{
\begin{table}[hbt]
\centering
\begin{tabular}{|l|c|c|}
\hline
Chapter                                                  & ISA? & Status \\
\hline
\hline
1.\ Introduction                                         & ---  & Frozen \\
2.\ Control and Status Registers (CSRs) Added to Harts   & Yes  & Frozen \\
3.\ Incoming MSI Controller (IMSIC)                      & Yes  & Frozen \\
4.\ Advanced Platform-Level Interrupt Controller (APLIC) & No   & Frozen \\
5.\ Interrupts for Machine and Supervisor Levels         & Yes  & Frozen \\
6.\ Interrupts for Virtual Machines (VS Level)           & Yes  & Frozen \\
7.\ Interprocessor Interrupts (IPIs)                     & No   & Frozen \\
8.\ IOMMU Support for MSIs to Virtual Machines           & No   & Frozen \\
\hline
\end{tabular}
\end{table}
}

An implementation adhering to the current document might not conform
to an eventual ratified Advanced Interrupt Architecture for {\RISCV}.
However, all chapters are \emph{frozen}, meaning they are not expected
to change significantly before being put up for ratification.

\section*{Changes for RC3 (Ratification Candidate 3)}

Removed the still-draft Duo-PLIC chapter to a separate document.

In Section~\ref{sec:virtIntrs-S} added the options
for bits 1 and 9 to be writable in CSR \z{mvien},
and specified the effects of setting each of these bits.

Upgraded Chapter~\ref{ch:IOMMU} (``IOMMU Support'')
to the \emph{frozen} state.

\section*{Changes for RC2}

The only normative changes to this document
between versions RC1 and RC2 were to clarify that
field IID of CSR \z{hvictl} must support all
unsigned integer values of the number of bits implemented
for that field, and that writes to \z{hvictl}
always set IID in the most straightforward way.

A comment was added to Chapter~\ref{ch:IPIs} warning about
the possible need for FENCE instructions when IPIs are
sent to other harts by writing MSIs to those harts' IMSICs.